### 2008-08-07 doc/backend_interface/Backends-arc1.tex
Updated the document according to the KnowARC review

\subsection{ARC1}
As of ARC version 0.9, which is to be the next generation ARC, the scripts are located in different directories
within the NorduGrid subversion tree~\cite{svn}. 
The job control interface scripts can be found under 
\begin{quote}
\texttt{arc1/trunk/src/services/a-rex/lrms/}
\end{quote}
while the information collectors are located in
\begin{quote}
\texttt{arc1/trunk/src/services/a-rex/infoproviders/}.
\end{quote}
The next generation of ARC supports both the classic NorduGrid information schema~\cite{is} and
currently a minimal set of the GLUE specification version 2.0 schema~\cite{glue2}.

@@

@@

If a grid user cancels his job, the message will reach the grid-manager.
The manager will then call the cancel-LRMS-job for the suitable back-end.
The cancel script is called with a text file containing information
about the job. It is called the GRAMi file. It contains such information
as the original xRSL option and the job id in the local LRMS. Cancel-LRMS-job
must then use that information to find the job and remove it from
the queue or actually cancel it if it is running in the LRMS.

The submit program is the most involved. It is called by the GM once
a new job arrives and needs to be submitted to the LRMS. Like cancel-LRMS-job
it is given the GRAMi file as argument on execution. Submit-LRMS-job
then has to set up the session directories, run-time environment and
anything else needed. Then it submits the job to the local LRMS.
This is normally done by generating a native job script for the LRMS
and then running the local submit command, but it can also be done
through an API if the LRMS supports it.

@@

Scan-LRMS-job is run periodically. Its job is to scan the LRMS for
jobs that have finished. Once it has found a finished job it will
write the exit-code of that job to the file job.\{gridid\}.lrms\_done
in the ARC jobstatus directory. Then it will call the gm-kick program
to notify GM about the finished job. Subsequently, the GM starts finalizing
the job.

Generally, two approaches are taken to find jobs that are done. One
is to ask the LRMS. First all started grid jobs are found in the jobstatus
directory%
\footnote{normally /var/spool/nordugrid/jobstatus/, but can be set via the controldir
variable of arc.conf%
}. This can be done by checking if the status is "INLRMS" in the jobs
status file%
\footnote{job.\{gridid\}.status%
}. Then the LRMS is asked for the status of those jobs. If they are
done they are marked as such and the GM is activated. The problem
with this approach is that for most LRMSs the information about finished
jobs are only available for a short period after the job finished.
Therefore appropriate steps have to be taken if the job is known to
have been started but is no longer present in the LRMS. The normal
approach is to analyze the jobs status output in the session directory.

The second approach is to parse the LRMSs log files. This method has
some drawbacks, e.g. the GM has to be allowed read access to the logs.
The back-end will then have to remember where in the log it was last
time it ran. This information will have to be stored in a file somewhere
on the front-end.

@@

\begin{table}
\caption{The \texttt{queue\_info} hash}
\label{t:queue_info}
\begin{center}
\begin{tabular}{|c|p{12cm}|}
\hline
key & value\\
\hline \hline
status & available slots in queue, negative number signals error\\
\hline
maxrunning & limit on number of running jobs\\
\hline
maxqueuable & limit on number of jobs queued on this queue\\
\hline
maxuserrun & limit on number of running jobs per user\\
\hline
maxcputime & limit on maximum CPU time for a job in this queue in minutes\\
\hline
mincputime & limit on minimum CPU time for a job in this queue in minutes\\
\hline
efaultcput & default CPU time limit for a job in this queue in minutes\\
\hline
maxwalltime & limit on maximum wall time for a job in this queue in minutes\\
\hline
minwalltime & limit on minimum wall time for a job in this queue in minutes\\
\hline
defaultwallt & default wall time limit for a job in this queue in minutes\\
\hline
running & number of CPUs used by running jobs (both grid and non-grid) in the
queue\\
\hline
queued & number of CPUs requested by queuing jobs in the queue\\
\hline
totalcpus & number of CPUs available to the queue\\
\hline
users & a hash with keys being local UNIX user names and values being
\texttt{users\_info} hashes as described in table \ref{t:users_info} \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{table}
\caption{The \texttt{jobs\_info} hash}
\label{t:jobs_info}
\begin{center}
\begin{tabular}{|c|p{12cm}|}
\hline
key & value\\
\hline \hline
status & LRMS jobstatus mapping\footnotemark[1]:
Running $\Rightarrow$ 'R',
Queued $\Rightarrow$ 'Q',
Suspended $\Rightarrow$ 'S',
Exiting $\Rightarrow$ 'E',
Other $\Rightarrow$ 'O'\\
\hline
rank & the job's position in the LRMS queue\\
\hline
mem & the memory usage of the job in kBs\\
\hline
walltime & consumed wall-time in minutes\\
\hline
cputime & consumed CPU-time in minutes\\
\hline
reqwalltime & wall-time requested by job in minutes\\
\hline
reqcputime & CPU-time requested by job in minutes\\
\hline
nodes & list of execution hosts\\
\hline
comment & array of string comments about the job in LRMS, can be an empty
array\\
\hline
\end{tabular}
\footnotetext[1]{LRMS job states mapping is described in\cite{is}}
\end{center}
\end{table}

@@

\begin{enumerate}
\item The computing nodes MUST be declared as cluster nodes (job-exclusive),
at the moment time-shared nodes are not supported by the ARC setup.
If you intend to run more than one job on a single processor then
you can use the virtual processor feature of PBS. 
\item For each queue, you MUST set one of the max\_user\_run or max\_running
attributes and its value SHOULD BE IN AGREEMENT with the number of
available resources (i.e. don't set the max\_running = 10 if you have
only six (virtual) processors in your system). If you set both max\_running
and max\_user\_run then obviously max\_user\_run has to be less or equal
to max\_running.
\item For the time being, do NOT set server limits like max\_running, please
use queue-based limits instead. 
\item Avoid using the max\_load and the ideal\_load directives. The node's
mom configuration file (<PBS home on the node>/mom\_priv/config) should not
contain any max\_load or ideal\_load directives. PBS closes down a
node (no jobs are allocated to it) when the load on the node reaches
the max\_load value. The max\_load value is meant for controlling
time-shared nodes. In case of job-exclusive nodes there is no need
for setting these directives, moreover incorrectly set values can
close down your node. 
\item Routing queues are not supported, those cannot be used within ARC.
\end{enumerate}
Additional useful configuration hints:

@@

\begin{itemize}
\item The PBS batch system back-end is enabled via setting the \texttt{lrms=\char`\"{}pbs\char`\"{}}
in the \texttt{{[}common]} configuration block. No need to specify
the flavour or the version number of the PBS, simply use the \texttt{{}``pbs''}
keyword as LRMS configuration value.
\item \texttt{pbs\_bin\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path to the qstat,pbsnodes,qmgr etc PBS
binaries.
\item \texttt{pbs\_log\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path of the PBS server logfiles which are
used by the GM to determine whether a PBS job is completed.
If not specified, the GM will use the \texttt{qstat} command
to find completed jobs. 
\item \texttt{lrmsconfig} from the \texttt{{[}cluster]} block can be used
as an optional free text field to describe further details about the
PBS configuration (e.g. \texttt{lrmsconfig=\char`\"{}single job per
processor\char`\"{}}).
\item \texttt{dedicated\_node\_string} from the \texttt{{[}cluster]} block
specifies the string which is used in the PBS node config to distinguish
the grid nodes from the rest. Suppose only a subset of nodes are available
for grid jobs, and these nodes have a common \texttt{node property}
string, this case the \texttt{dedicated\_node\_string} should be set
to this value and only the nodes with the corresponding PBS \texttt{node
property} are counted as grid enabled nodes. Setting the \texttt{dedicated\_node\_string}
to the value of the PBS \texttt{node property} of the grid-enabled
nodes will influence how the totalcpus, user freecpus is calculated.
No need to set this attribute if the cluster is fully available for
the grid and the PBS configuration does not use the \texttt{node property}
method to assign certain nodes to grid queues.
\item \texttt{{[}queue/queuename]} block. For each grid-enabled (or grid
visible) PBS queue a corresponding \texttt{{[}queue]} block must be
defined. \texttt{queuename} should be the PBS queue name.
\item \texttt{scheduling\_policy} from the \texttt{{[}queue/queuename]}
block describes the scheduling policy of the queue. PBS by default
offers the FIFO scheduler, many sites run the MAUI. At the moment
\texttt{FIFO} \& \texttt{MAUI} are supported values. If you have a
MAUI scheduler you should specify the \char`\"{}MAUI\char`\"{} value
since it modifies the way the queue resources are calculated. By default
the \char`\"{}FIFO\char`\"{} scheduler type is assumed.
\item \texttt{maui\_bin\_path} from the \texttt{{[}queue/queuename]} block
sets the path of the maui commands like \texttt{showbf} when \char`\"{}MAUI\char`\"{}
is specified as \texttt{scheduling\_policy} value. This parameter
can be set in the \texttt{{[}common]} block as well. 
\item \texttt{queue\_node\_string} of the \texttt{{[}queue/queuename]} block
can be used similar to the \texttt{dedicated\_node\_string}. In PBS
you can assign nodes to a queue (or a queue to nodes) by using the
\texttt{node property} PBS node configuration method and assigning
the marked nodes to the queue (setting the \texttt{resources\_default.neednodes}
= \texttt{queue\_node\_string} for that queue). This parameter should
contain the \texttt{node property} string of the queue-assigned nodes.
Setting the \texttt{queue\_node\_string} changes how the queue-totalcpus,
user freecpus are determined for this queue. 
\end{itemize}

@@

Sun Grid Engine, formerly known as Codine, is an open source batch
system maintained by Sun. It runs on Linux, and Solaris.

@@

\begin{tabular}{|l|p{12cm}|}
\hline
Property & Comments \\
\hline \hline
BaseType & \texttt{Share} \\
\hline
Name & from the configuration \\
\hline
Description & ``comment'' configuration option \\
\hline
MappingQueue & from the configuration \\
\hline
SchedulingPolicy & if maui is used, set to \texttt{fairshare} \\
\hline
ServingState & has the same value as ComputingEndpoint.ServingState \\
\hline
\end{tabular}

>>>>

\subsection{ARC1}
As of ARC version 0.9, which is to be the next generation ARC, the scripts are located in different directories
within the NorduGrid subversion tree~\cite{svn}. 
The job control interface scripts can be found under 
\begin{quote}
\texttt{arc1/trunk/src/services/a-rex/lrms/}
\end{quote}
while the information collectors are located in
\begin{quote}
\texttt{arc1/trunk/src/services/a-rex/infoproviders/}.
\end{quote}

The backend scripts install along with the ARC1 code.

The next generation of ARC supports both the classic NorduGrid information schema~\cite{is} and
currently a minimal set of the GLUE specification version 2.0 schema~\cite{glue2}.

@@

The submit program is the most involved. It is called by the GM once
a new job arrives and needs to be submitted to the LRMS. It is given
the GRAMi file as argument on execution. The GRAMi file is a file in
the job control directory containing the job description in a flat list
of key-value pairs. This file is created by GM and is based on the JSDL
job description. Submit-LRMS-job then has to set up the session
directories, run-time environment and anything else needed. Then it
submits the job to the local LRMS. This is normally done by generating
a native job script for the LRMS and then running the local submit
command, but it can also be done through an API if the LRMS supports it.

@@

If a grid user cancels his job, the message will reach the grid-manager.
The manager will then call the cancel-LRMS-job for the suitable back-end.
The cancel script is called with the GRAMi file containing information
about the job such as the job id in the LRMS. Cancel-LRMS-job must then
use that information to find the job and remove it from the queue or
actually cancel it if it is running in the LRMS.

@@

The scan-LRMS-job is run periodically. Its job is to scan the LRMS for
jobs that have finished. Once it has found a finished job it will
write the exit-code of that job to the file job.\{gridid\}.lrms\_done
in the ARC job status directory\footnote{normally /var/spool/nordugrid/jobstatus/,
but can be set via the controldir variable of arc.conf}. Then it will call the
gm-kick program to notify GM about the finished job. Subsequently, the GM starts
finalizing the job.

Generally, two approaches are taken to find jobs which are finished in LRMS. One
is to directly ask the LRMS. Since all started grid jobs have its own status
file\footnote{job.\{gridid\}.status} found in the job status directory,
this can be done by checking if the status is "INLRMS" in this file. If so, a
call to the LRMS is made asking for the status of the job (or jobs if several
jobs have status "INLRMS"). If it is finished, it is marked as such in the
job status directory, and the gm-kick program is activated. For most LRMSs the
information about finished jobs are only available for a short period of time after
the job finished. Therefore appropriate steps have to be taken if the job has the
status "INLRMS" in the job status directory, but is no longer present in the LRMS.
The normal approach is to analyze the job's status output in the session directory.

The second approach is to parse the LRMSs log files. This method has
some drawbacks like e.g.: the GM has to be allowed read access to the logs.
The back-end will then have to remember where in the log it was last
time it ran. This information will have to be stored in a file somewhere
on the front-end.

@@

\begin{table}
\caption{The \texttt{queue\_info} hash}
\label{t:queue_info}
\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{|c|p{12cm}|}
\hline
key & value\\
\hline \hline
status & available slots in queue, negative number signals error\\
\hline
maxrunning & limit on number of running jobs\\
\hline
maxqueuable & limit on number of jobs queued on this queue\\
\hline
maxuserrun & limit on number of running jobs per user\\
\hline
maxcputime & limit on maximum CPU time for a job in this queue in minutes\\
\hline
mincputime & limit on minimum CPU time for a job in this queue in minutes\\
\hline
efaultcput & default CPU time limit for a job in this queue in minutes\\
\hline
maxwalltime & limit on maximum wall time\footnote{Also known as Wall clock
time, \url{http://en.wikipedia.org/wiki/Wall_clock_time}} for a job in this queue in minutes\\
\hline
minwalltime & limit on minimum wall time for a job in this queue in minutes\\
\hline
defaultwallt & default wall time limit for a job in this queue in minutes\\
\hline
running & number of CPUs used by running jobs (both grid and non-grid) in the
queue\\
\hline
queued & number of CPUs requested by queuing jobs in the queue\\
\hline
totalcpus & number of CPUs available to the queue\\
\hline
users & a hash with keys being local UNIX user names and values being
\texttt{users\_info} hashes as described in table \ref{t:users_info} \\
\hline
\end{tabular}
\end{center}
\end{minipage}
\end{table}

@@

\begin{table}
\caption{The \texttt{jobs\_info} hash}
\label{t:jobs_info}
\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{|c|p{12cm}|}
\hline
key & value\\
\hline \hline
status & LRMS jobstatus mapping\footnote{LRMS job states mapping is described in\cite{is}}:
Running $\Rightarrow$ 'R',
Queued $\Rightarrow$ 'Q',
Suspended $\Rightarrow$ 'S',
Exiting $\Rightarrow$ 'E',
Other $\Rightarrow$ 'O'\\
\hline
rank & the job's position in the LRMS queue\\
\hline
mem & the memory usage of the job in kBs\\
\hline
walltime & consumed wall time in minutes\\
\hline
cputime & consumed CPU-time in minutes\\
\hline
reqwalltime & wall time requested by job in minutes\\
\hline
reqcputime & CPU-time requested by job in minutes\\
\hline
nodes & list of execution hosts\\
\hline
comment & array of string comments about the job in LRMS, can be an empty
array\\
\hline
\end{tabular}
\end{center}
\end{minipage}
\end{table}

@@

\begin{enumerate}
\item The computing nodes MUST be declared as cluster nodes (job-exclusive),
at the moment time-shared nodes are not supported by the ARC setup.
If you intend to run more than one job on a single processor then
you can use the virtual processor feature of PBS. 
\item For each queue, you MUST set one of the max\_user\_run or max\_running
attributes and its value SHOULD BE IN AGREEMENT with the number of
available resources (i.e. don't set the max\_running = 10 if you have
only six (virtual) processors in your system). If you set both max\_running
and max\_user\_run then obviously max\_user\_run has to be less or equal
to max\_running.
\item For the time being, do NOT set server limits like max\_running, please
use queue-based limits instead. 
\item Avoid using the max\_load and the ideal\_load directives. The Node Manager
(MOM) configuration file (<PBS home on the node>/mom\_priv/config) should not
contain any max\_load or ideal\_load directives. PBS closes down a
node (no jobs are allocated to it) when the load on the node reaches
the max\_load value. The max\_load value is meant for controlling
time-shared nodes. In case of job-exclusive nodes there is no need
for setting these directives, moreover incorrectly set values can
close down your node. 
\item Routing queues are not supported, those cannot be used within ARC.
\end{enumerate}
Additional useful configuration hints:

@@

\begin{itemize}
\item The PBS batch system back-end is enabled via setting the \texttt{lrms=\char`\"{}pbs\char`\"{}}
in the \texttt{{[}common]} configuration block. No need to specify
the flavour or the version number of the PBS, simply use the \texttt{{}``pbs''}
keyword as LRMS configuration value.
\item \texttt{pbs\_bin\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path to the qstat,pbsnodes,qmgr etc PBS
binaries.
\item \texttt{pbs\_log\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path of the PBS server logfiles which are
used by the GM to determine whether a PBS job is completed.
If not specified, the GM will use the \texttt{qstat} command
to find completed jobs. 
\item \texttt{lrmsconfig} from the \texttt{{[}cluster]} block can be used
as an optional free text field to describe further details about the
PBS configuration (e.g. \texttt{lrmsconfig=\char`\"{}single job per
processor\char`\"{}}).
\item \texttt{dedicated\_node\_string} from the \texttt{{[}cluster]} block
specifies the string which is used in the PBS node config to distinguish
the grid nodes from the rest. Suppose only a subset of nodes are available
for grid jobs, and these nodes have a common \texttt{node property}
string, this case the \texttt{dedicated\_node\_string} should be set
to this value and only the nodes with the corresponding PBS \texttt{node
property} are counted as grid enabled nodes. Setting the \texttt{dedicated\_node\_string}
to the value of the PBS \texttt{node property} of the grid-enabled
nodes will influence how the totalcpus, user freecpus is calculated.
No need to set this attribute if the cluster is fully available for
the grid and the PBS configuration does not use the \texttt{node property}
method to assign certain nodes to grid queues.
\item \texttt{{[}queue/queuename]} block. For each grid-enabled (or grid
visible) PBS queue a corresponding \texttt{{[}queue]} block must be
defined. \texttt{queuename} should be the PBS queue name.
\item \texttt{scheduling\_policy} from the \texttt{{[}queue/queuename]}
block describes the scheduling policy of the queue. PBS by default
offers the FIFO scheduler, many sites run the MAUI. At the moment
\texttt{FIFO} \& \texttt{MAUI} are supported values. If you have a
MAUI scheduler you should specify the \char`\"{}MAUI\char`\"{} value
since it modifies the way the queue resources are calculated. By default
the \char`\"{}FIFO\char`\"{} scheduler type is assumed.
\item \texttt{maui\_bin\_path} from the \texttt{{[}queue/queuename]} block
sets the path of the MAUI commands like \texttt{showbf} when \char`\"{}MAUI\char`\"{}
is specified as \texttt{scheduling\_policy} value. This parameter
can be set in the \texttt{{[}common]} block as well. 
\item \texttt{queue\_node\_string} of the \texttt{{[}queue/queuename]} block
can be used similar to the \texttt{dedicated\_node\_string}. In PBS
you can assign nodes to a queue (or a queue to nodes) by using the
\texttt{node property} PBS node configuration method and assigning
the marked nodes to the queue (setting the \texttt{resources\_default.neednodes}
= \texttt{queue\_node\_string} for that queue). This parameter should
contain the \texttt{node property} string of the queue-assigned nodes.
Setting the \texttt{queue\_node\_string} changes how the queue-totalcpus,
user freecpus are determined for this queue. 
\end{itemize}

@@

Sun Grid Engine, formerly known as Codine, is an open source batch
system maintained by Sun. It is supported on Linux, and Solaris in
addition to a numerous other systems.

@@

\begin{tabular}{|l|p{12cm}|}
\hline
Property & Comments \\
\hline \hline
BaseType & \texttt{Share} \\
\hline
Name & from the configuration \\
\hline
Description & ``comment'' configuration option \\
\hline
MappingQueue & from the configuration \\
\hline
SchedulingPolicy & if MAUI is used, set to \texttt{fairshare} \\
\hline
ServingState & has the same value as ComputingEndpoint.ServingState \\
\hline
\end{tabular}
